---
title: "Class Exercise #2"
author: "Nicola Casarin & Francesco Catalfamo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this short paper, we expose the workflow we built in order to complete the second class exercise of the Data Access and Regulation class.
Our goal was to implement the learned scraping techniques in a real-life scenario. The selected scraping target were the posts published during 2016 onto [Beppe Grillo's blog](https://beppegrillo.it/).

The team is composed of [Nicola Casarin](https://github.com/n-oise) and [Francesco Catalfamo](https://github.com/FCatalfamo).

## Tasks and Resolutions

#### Preparation

For our work we used six main packages:

```{r Preparation, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(curl)
library(here)
library(stringr)
```

Furthermore, in order to simplify the workflow we defined two functions: get_page() and page_name().

The first function scrape sites in a polite way and takes six arguments. *url*, a character vector with length 1 or more containing the links; *dest*, for defining the downloading path; *my_email*, for defining the scraper mail; *agent*, a logical if **TRUE** return the User Agent; *filetype*, for defining the output file type; lastly the *scrapeText*, a logical if **TRUE** scrape the download HTML content and assign it to an object.

```{r get_page}

get_page <- function(url, dest = "", my_email = "", agent = F, filetype = ".html", scrapeText = F) { 
  # Donwload file from a list of URLs controlling for http status code. 
    stopifnot(is.logical(agent))
    stopifnot(is.logical(scrapeText))
    UA <- ifelse(agent == FALSE, "", R.Version()$version.string)
    
    
    i <- 0
    for (i in 1:length(url)) {
      url_step <- url[i]
      httpReq <- httr::GET(url_step,
                     httr::add_headers(
                       From = my_email, 
                       `User-Agent` = UA
                       ))
      code <- httr::status_code(httpReq)
      
      
      # If status code is OK then:
      # download file with the filename and path as in page_name 
      if (code == 200) {
         bin <- httr::content(httpReq,
                        as = "raw")
         writeBin(object = bin,
                  con = page_name(url = url_step, dest, filetype))
         #Scrape the main text if specified 
         if (scrapeText == T | code == 200) {
           assign(basename(url_step), 
                  read_html(httpReq), 
                  envir = parent.frame())
         }
        
      } else if (code == 404) { # If status code is Not Found
        cat("Bad luck. Error ",
            code,
            " Resource not found! :(",
            sep = "")
       } else { # For every other status code report a generic error.
         cat("Error ",
             code,
             ". Check your URL!",
             sep = "")
       }
      
      # Sleep every round!
      Sys.sleep(2)
      
    }
}
```

The latter function is used inside the *get_page()* function and permits to defining the file name from the hostname and the basename. The function takes three arguments defined in the main *get_page()* function.

```{r page_name()}
page_name <- function(url, dest = "", filetype = ".html") {
  if (is.character(url)) {
    name <- stringr::str_extract(url, "[^https?://][^/]*")
    filename <- paste0("/",
                       name,
                       "_",
                       basename(url),
                       filetype)
    if (dest == "") { 
      #if the dest param is empty page_name define  path at the project's root
      if (name == basename(url)) {
      path <- paste0(here::here(),
                     "/",
                     basename(url),
                     filetype)
      } 
      else {
        path <- paste0(here::here(),
                       filename)
      } 
      return(path)
    } 
    else if (dest != ""){ 
      #if the dest param is NOT empty the function define 
      #the path to the project's subdirectory as in dest
      if (name == basename(url)) {
        path <- paste0(here::here(dest),
                       "/",
                       basename(url),
                       filetype)
      } 
      else {
        path <- paste0(here::here(dest),
                       filename)
      } 
      return(path)
    }
  } 
  else{
    cat("Not a string!")
   }
}
```

#### Task #1: Inspect the Robots.txt


#### Task #2: Download the page in a polite way


#### Task #3: Create a dataframe with all the links in the downloaded page

#### Task #4: Download and scrape all the 2016 posts

#### Task #5: Sketch a spider scraper

## Contributions

